---
title: "Initial Variable Exploratory"
author: "Kristine Dinh"
date: "4/17/2020"
output: html_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

```


```{r load_libraries}

library(tidyverse)
library(magrittr)
library(broom)
library(kableExtra)
library(knitr)

```


```{r load_data}

setwd(here::here())

source("R/data.R")

```


```{r function}

Explore <- function(df, group, reorder = FALSE){
  
  print_list <- list()
  
  df %<>% select(dx_heart, xvar = group)
  
  df_group <- df %>% 
    group_by(xvar) %>% 
    summarise(count = n(), 
              prop = mean(dx_heart))
  
  ratio <- max(df_group$count) / max(df_group$prop)
  
  if(reorder == TRUE){
    plot <- ggplot(df_group, aes(x = reorder(xvar, -count), group = 1))
  } else if (reorder == FALSE) {
    plot <- ggplot(df_group, aes(x = xvar, group = 1))
  }
  
  
  plot <- plot + 
    geom_bar(aes(y = count), stat = "identity", col = "black", fill = "grey") + 
    geom_point(aes(y = prop * ratio), col = "red", size = 2) + 
    geom_line(aes(y = prop * ratio), col = "red", size = 1) + 
    scale_y_continuous(sec.axis = sec_axis(~./ratio, name = "Average prop of having heart disease")) +
    labs(title = paste("Distribution plot of", group), 
         subtitle = "With Average proportion of getting diagnosed with heart disease") + 
    labs(x = group) + 
    theme_minimal()
  
  distribution_table <- df_group %>%
    kable(digits = 4, caption = "distribution table") %>% 
    kable_styling("bordered", full_width = FALSE)
    
  fit <- glm(dx_heart ~ ., data = df, family = "binomial")
  
  tidy <- tidy(fit)
  
  confint <- confint(fit)
  
  summary_statistics <- cbind(tidy, confint) %>% 
    kable(row.names = FALSE, digits = 3, caption = "summary statistics") %>% 
    kable_styling("bordered", full_width = FALSE)
  
  print_list$plot <- plot
  print_list$distribution_table <- distribution_table
  print_list$summary_statistics <- summary_statistics
  
  return(print_list)
}

```


#### **Gender** 

```{r}

gender <- Explore(data, "gender", reorder = TRUE)

gender$plot

gender$distribution_table

gender$summary_statistics

```


#### **Age**

```{r}

age <- Explore(data, "age")

age$plot

age$summary_statistics

```


**Chest Pain Type**

```{r}

chest_pain_type <- Explore(data, "chest_pain_type")

chest_pain_type$plot

chest_pain_type$distribution_table

chest_pain_type$summary_statistics

```


#### **Resting Blood Pressure**

```{r}

rest_bp_bins <- Explore(data, "rest_bp_bins")

rest_bp_bins$plot

rest_bp <- Explore(data, "rest_bp")

rest_bp$summary_statistics

```


**Serum Cholestoral in mg/dl**

```{r}

chol_bins <- Explore(data, "chol_bins")

chol_bins$plot

chol <- Explore(data, "chol")

chol$summary_statistics

```


**Fasting Blood Sugar > 120 mg/dl**

```{r}

fbs <- Explore(data, "fbs", reorder = TRUE)

fbs$plot

fbs$distribution_table

fbs$summary_statistics

```


**Resting Electrocardiographic Results**

```{r}

rest_ecg <- Explore(data, "rest_ecg")

rest_ecg$plot

rest_ecg$distribution_table

rest_ecg$summary_statistics

```


**Maximum Heart Rate Achieved**

```{r}

max_heart_rate_bins <- Explore(data, "max_heart_rate_bins")

max_heart_rate_bins$plot

max_heart_rate <- Explore(data, "max_heart_rate")

max_heart_rate$summary_statistics

```


**Exercise Induced Angina**

```{r}

exercise_angina <- Explore(data, "exercise_angina", reorder = TRUE)

exercise_angina$plot

exercise_angina$distribution_table

exercise_angina$summary_statistics

```


**ST Depression Induced by Exercise Relative to Rest**

```{r}

oldpeak_bins <- Explore(data, "oldpeak_bins")

oldpeak_bins$plot

oldpeak <- Explore(data, "oldpeak")

oldpeak$summary_statistics

```


**The Slope of the Peak Exercise ST Segment**

```{r}

slope <- Explore(data, "slope")

slope$plot

slope$distribution_table

slope$summary_statistics

```


**Number of Major Vessels (0-3) Colored by Flourosopy**

```{r}

vessels <- Explore(data, "vessels")

vessels$plot

vessels$distribution_table

vessels$summary_statistics

```

**Thal: 3 = normal; 6 = fixed defect; 7 = reversable defect**

```{r}

thal <- Explore(data, "thal")

thal$plot

thal$distribution_table

thal$summary_statistics

```

























